---
title: "BIRCH Report"
author: "ACBRI Bioinformatics Team"
date: 'Report generated on `r format(Sys.time(), "%B %d, %Y")`'

output:
  html_document:
    anchor_sections: false
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
    fig_width: 11
    fig_height: 7
    code_folding: "hide"
    theme: united
toc-title: "Table of Contents"

params:
  input_plot1_reac : input_plot1_reac
  balloon_plot_reac_report : balloon_plot_reac_report
  pareto_plot_reac : pareto_plot_reac
  missing_plots2_reac_report : missing_plots2_reac_report
  missing_plot3_reac : missing_plot3_reac
  init_table_reac : init_table_reac
  filt_plot1_reac : filt_plot1_reac
  balloon_plot_reac_report2 : balloon_plot_reac_report2
  filt_plot2_reac : filt_plot2_reac
  filt_plot3_reac : filt_plot3_reac
  filtered_plot1_reac : filtered_plot1_reac
  filtered_plot2_reac : filtered_plot2_reac
  missing_plots2_reac_report2 : missing_plots2_reac_report2
  filtered_plot4_reac : filtered_plot4_reac
  unnorm_box_reac : unnorm_box_reac
  norm_box_reac : norm_box_reac
  pvca_before_bc_reac : pvca_before_bc_reac
  pvca_after_across_bc_reac : pvca_after_across_bc_reac
  pvca_after_bc_reac : pvca_after_bc_reac
  pca_before_bc_reac : pca_before_bc_reac
  pca_after_across_bc_reac : pca_after_across_bc_reac
  pca_after_bc_reac : pca_after_bc_reac
  hca_before_bc_reac : hca_before_bc_reac
  hca_after_across_bc_reac : hca_after_across_bc_reac
  hca_after_bc_reac : hca_after_bc_reac
  irt_before_bc_reac : irt_before_bc_reac
  irt_after_across_bc_reac : irt_after_across_bc_reac
  irt_after_bc_reac : irt_after_bc_reac
  corr_before_bc_reac : corr_before_bc_reac
  corr_after_across_bc_reac : corr_after_across_bc_reac
  corr_after_bc_reac : corr_after_bc_reac
---

```{r Load R Libraries, include=FALSE}
library(knitr)
```

```{js Move the top-right CODE button, echo=FALSE}
// https://stackoverflow.com/questions/43009788/insert-a-logo-in-upper-right-corner-of-r-markdown-html-document
$(document).ready(function() {
  $('#header').css('margin-right', '180px')
});
```

```{r, include = FALSE}
if (identical(rmarkdown::metadata$runtime, "shiny"))
  shiny::setProgress(0.25)  # set progress to 50%

htmltools::tagList(ggplotly(ggplot()))
htmltools::tagList(plotly::ggplotly(ggplot()))
```

## Initial analysis {.tabset}
### Sample distribution
```{r, include = TRUE}
grid.draw(params$input_plot1_reac)
```

### Sample matrix
```{r, include = TRUE, echo=FALSE, results="asis"}

for (i in 1:length(params$balloon_plot_reac_report)) {
  params$balloon_plot_reac_report[[i]]
}

#params$balloon_plot_reac_report[[1]]
#params$balloon_plot_reac_report[[2]]

```

### Missingness distribution
```{r}
params$pareto_plot_reac

for (i in 1:length(params$missing_plots2_reac_report)) {
  params$missing_plots2_reac_report[[i]]
}

#params$missing_plots2_reac_report[[1]]
#params$missing_plots2_reac_report[[2]]

params$missing_plot3_reac
```

### Take aways
```{r}
knitr::kable(params$init_table_reac)
```

## Filtering {.tabset}
### Fragments per Sample
```{r}
params$filt_plot1_reac
```

### Fragments per group
```{r}
params$filt_plot2_reac
```

### Fragments per batch
```{r}
params$filt_plot3_reac
```

## Filtered results {.tabset}
### Sample distribution after filtering
```{r, include = TRUE}
grid.draw(params$filtered_plot1_reac)
```

### Sample matrix after filtering
```{r, include = TRUE}

for (i in 1:length(params$balloon_plot_reac_report2)) {
  params$balloon_plot_reac_report2[[i]]
}

#params$balloon_plot_reac_report2[[1]]
#params$balloon_plot_reac_report2[[2]]

```

### Missingness after filtering
```{r}
params$filtered_plot2_reac

for (i in 1:length(params$missing_plots2_reac_report2)) {
  params$missing_plots2_reac_report2[[i]]
}

#params$missing_plots2_reac_report2[[1]]
#params$missing_plots2_reac_report2[[2]]

params$filtered_plot4_reac
```

### Normalization
```{r}
params$unnorm_box_reac
params$norm_box_reac

if (identical(rmarkdown::metadata$runtime, "shiny"))
  shiny::setProgress(0.5)  # set progress to 100%
```

## Batch corrected results 
## PVCA {.tabset}
### Before batch correction
```{r}
params$pvca_before_bc_reac
```

### After combat only correction
```{r}
params$pvca_after_across_bc_reac
```

### After combat+loess correction
```{r}
params$pvca_after_bc_reac
```

## PCA {.tabset}
### Before batch correction
```{r, include = TRUE}
grid.draw(params$pca_before_bc_reac)
```

### After combat only correction
```{r, include = TRUE}
grid.draw(params$pca_after_across_bc_reac)
```

### After combat+loess correction
```{r, include = TRUE}
grid.draw(params$pca_after_bc_reac)
```

## iRT mapping {.tabset}
### Before batch correction
```{r}
params$irt_before_bc_reac
```

### After combat only correction
```{r}
params$irt_after_across_bc_reac
```

### After combat+loess correction
```{r}
params$irt_after_bc_reac
```

## Correlation {.tabset}
### Before batch correction
```{r}
params$corr_before_bc_reac
```

### After combat only correction
```{r}
params$corr_after_across_bc_reac
```

### After combat+loess correction
```{r}
params$corr_after_bc_reac

if (identical(rmarkdown::metadata$runtime, "shiny"))
  shiny::setProgress(1)  # set progress to 100%
```