---
title: "BIRCH Report"
author: "ACBRI Bioinformatics Team"
date: 'Report generated on `r format(Sys.time(), "%B %d, %Y")`'

output:
  html_document:
    anchor_sections: false
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
    fig_width: 11
    fig_height: 7
    code_folding: "hide"
    theme: united
toc-title: "Table of Contents"

params:
  sample_matrix : sample_matrix
  sample_matrix_filt : sample_matrix_filt
  joined_df : joined_df
  joined_df2 : joined_df2
  input_plot1_reac : input_plot1_reac
  balloon_plot_reac_report : balloon_plot_reac_report
  pareto_plot_reac : pareto_plot_reac
  missing_plots2_reac_report : missing_plots2_reac_report
  missing_plot3_reac : missing_plot3_reac
  init_table_reac : init_table_reac
  filt_plot1_reac : filt_plot1_reac
  balloon_plot_reac_report2 : balloon_plot_reac_report2
  filt_plot2_reac : filt_plot2_reac
  filt_plot3_reac : filt_plot3_reac
  filtered_plot1_reac : filtered_plot1_reac
  filtered_plot2_reac : filtered_plot2_reac
  missing_plots2_reac_report2 : missing_plots2_reac_report2
  filtered_plot4_reac : filtered_plot4_reac
  init_table_reac2 : init_table_reac2
  unnorm_box_reac : unnorm_box_reac
  norm_box_reac : norm_box_reac
  pvca_before_bc_reac : pvca_before_bc_reac
  pvca_after_across_bc_reac : pvca_after_across_bc_reac
  pvca_after_bc_reac : pvca_after_bc_reac
  pca_before_bc_reac : pca_before_bc_reac
  pca_after_across_bc_reac : pca_after_across_bc_reac
  pca_after_bc_reac : pca_after_bc_reac
  hca_before_bc_reac : hca_before_bc_reac
  hca_after_across_bc_reac : hca_after_across_bc_reac
  hca_after_bc_reac : hca_after_bc_reac
  irt_before_bc_reac : irt_before_bc_reac
  irt_after_across_bc_reac : irt_after_across_bc_reac
  irt_after_bc_reac : irt_after_bc_reac
  corr_before_bc_reac : corr_before_bc_reac
  corr_after_across_bc_reac : corr_after_across_bc_reac
  corr_after_bc_reac : corr_after_bc_reac
---

```{r Load R Libraries, include=FALSE}
library(knitr)
```

```{js Move the top-right CODE button, echo=FALSE}
// https://stackoverflow.com/questions/43009788/insert-a-logo-in-upper-right-corner-of-r-markdown-html-document
$(document).ready(function() {
  $('#header').css('margin-right', '180px')
});
```

---
## Introduction
```{r ACBRI introduction, results='asis', echo=FALSE}
cat("\n\n
Hello and welcome! We are pleased to be trusted with your data and present to you a visual summary of your dataset.

Batch effect correction is the procedure of removing variability from your data that is not due to your variable of interest. Batch effects are due to technical differences between your samples, such as the type of instrument or even the technician that ran the sample.

For any concerns about or suggestions to improve this report, let the team know at **GroupHeartBioinformaticsSupport@cshs.org**, and we will consider your feedback. We are happy to support your computational needs.
\n\n", sep="")
```

```{r, include = FALSE}
if (identical(rmarkdown::metadata$runtime, "shiny"))
  shiny::setProgress(0.25)  # set progress to 50%

htmltools::tagList(ggplotly(ggplot()))
htmltools::tagList(plotly::ggplotly(ggplot()))
```

## Initial analysis {.tabset}
### Sample distribution
```{r, include = TRUE, echo=F, results='asis', message=FALSE, warning=FALSE, fig.width=16, fig.height=8}
grid.draw(params$input_plot1_reac)
```

### Sample matrix
```{r, include = TRUE, echo=F, results='asis', message=FALSE, warning=FALSE}

if (length(params$balloon_plot_reac_report) == 1) {
  params$balloon_plot_reac_report[[1]]
} else {
  for (i in params$balloon_plot_reac_report) {
    print(i)
  }
}

#params$balloon_plot_reac_report[[1]]
#params$balloon_plot_reac_report[[2]]

```

### Missingness distribution
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$pareto_plot_reac

if (length(params$missing_plots2_reac_report) == 1) {
  params$missing_plots2_reac_report[[1]]
} else {
  for (i in params$missing_plots2_reac_report) {
    print(i)
  }
}

#params$missing_plots2_reac_report[[1]]
#params$missing_plots2_reac_report[[2]]

params$missing_plot3_reac
```

### Take aways
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
knitr::kable(params$init_table_reac)

cat("\n\n Count of missingness within your entire dataset (i.e., cells in the matrix with NAs) is ", toString(sum(params$joined_df$value)), "and the percent of missing data is ", toString(round((sum(params$joined_df$value)/(nrow(params$sample_matrix)*(ncol(params$sample_matrix)-1)))*100, 2)), "%. This percentage of missingness in overall data should also be < 50% for effective batch correction.", sep=" ")

```

## Filtering {.tabset}
### Fragments per Sample
```{r echo=F, results='asis', message=FALSE, warning=FALSE}
params$filt_plot1_reac
```

### Fragments per group
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$filt_plot2_reac
```

### Fragments per batch
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$filt_plot3_reac
```

## Filtered results {.tabset}
### Filtered stats
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
cat("\n\n __Number of samples:__")
cat("\n\n Initial number of samples was: ", toString(ncol(params$sample_matrix)-1), sep=" ")
cat("\n\n Number of samples remaining after filtering are: ", toString(ncol(params$sample_matrix_filt)-1), sep=" ")
    
cat("\n\n __Number of fragments:__")
cat("\n\n Initial number of fragments was: ", toString(nrow(params$sample_matrix)), sep=" ")
cat("\n\n Number of fragments remaining after filtering are: ", toString(nrow(params$sample_matrix_filt)), sep=" ")

cat("\n\n __Missingness before/after filtering:__")
cat("\n\n Count of missingness within your entire dataset (i.e., cells in the matrix with NAs) was ", toString(sum(params$joined_df$value)), "and the percent of missing data was ", toString(round((sum(params$joined_df$value)/(nrow(params$sample_matrix)*(ncol(params$sample_matrix)-1)))*100, 2)), "%. This percentage of missingness in overall data should also be < 50% for effective batch correction.", sep=" ")

cat("\n\n Count of missingness within your entire dataset (i.e., cells in the matrix with NAs) after filtering is ", toString(sum(params$joined_df2$value)), "and the percent of missing data is ", toString(round((sum(params$joined_df2$value)/(nrow(params$sample_matrix_filt)*(ncol(params$sample_matrix_filt)-1)))*100, 2)), "%. This percentage of missingness in overall data should also be < 50% for effective batch correction.", sep="")
```

### Sample distribution after filtering
```{r, include = TRUE, echo=F, results='asis', message=FALSE, warning=FALSE, fig.width=16, fig.height=8}
grid.draw(params$filtered_plot1_reac)
```

### Sample matrix after filtering
```{r, include = TRUE, echo=F, results='asis', message=FALSE, warning=FALSE}

if (length(params$balloon_plot_reac_report2) == 1) {
  params$balloon_plot_reac_report2[[1]]
} else {
  for (i in params$balloon_plot_reac_report2) {
    print(i)
  }
}


#params$balloon_plot_reac_report2[[1]]
#params$balloon_plot_reac_report2[[2]]

```

### Missingness after filtering
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$filtered_plot2_reac

if (length(params$missing_plots2_reac_report2) == 1) {
  params$missing_plots2_reac_report2[[1]]
} else {
  for (i in params$missing_plots2_reac_report2) {
    print(i)
  }
}

#params$missing_plots2_reac_report2[[1]]
#params$missing_plots2_reac_report2[[2]]

params$filtered_plot4_reac
```

### Take aways
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
knitr::kable(params$init_table_reac2)

cat("\n\n Count of missingness within your entire dataset (i.e., cells in the matrix with NAs) is ", toString(sum(params$joined_df2$value)), "and the percent of missing data is ", toString(round((sum(params$joined_df2$value)/(nrow(params$sample_matrix_filt)*(ncol(params$sample_matrix_filt)-1)))*100, 2)), "%. This percentage of missingness in overall data should also be < 50% for effective batch correction.", sep=" ")
```

## Normalization {.tabset}
### Before normalization
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$unnorm_box_reac
```

### After normalization
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$norm_box_reac

if (identical(rmarkdown::metadata$runtime, "shiny"))
  shiny::setProgress(0.5)  # set progress to 100%
```

## Batch corrected results 
## PVCA {.tabset}
### Before batch correction
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$pvca_before_bc_reac[[2]]
```

### After combat only correction
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$pvca_after_across_bc_reac[[2]]
```

### After combat+loess correction
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$pvca_after_bc_reac[[2]]
```

## PCA {.tabset}
### Before batch correction
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
grid.draw(params$pca_before_bc_reac)
```

### After combat only correction
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
grid.draw(params$pca_after_across_bc_reac)
```

### After combat+loess correction
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
grid.draw(params$pca_after_bc_reac)
```

## iRT mapping {.tabset}
### Before batch correction
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$irt_before_bc_reac
```

### After combat only correction
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$irt_after_across_bc_reac
```

### After combat+loess correction
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$irt_after_bc_reac
```

## Correlation {.tabset}
### Before batch correction
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$corr_before_bc_reac
```

### After combat only correction
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$corr_after_across_bc_reac
```

### After combat+loess correction
```{r, echo=F, results='asis', message=FALSE, warning=FALSE}
params$corr_after_bc_reac

if (identical(rmarkdown::metadata$runtime, "shiny"))
  shiny::setProgress(1)  # set progress to 100%
```